---
title: "LCMS data preprocessing"
output: html_notebook
---
# DATA IMPORT
```{r}
library(xcms)
library(faahKO)
library(RColorBrewer)
library(pander)
library(magrittr)
```

```{r}
## Get the full path to the CDF files
cdfs <- dir(system.file("cdf", package = "faahKO"), full.names = TRUE,
        recursive = TRUE)
## Create a phenodata data.frame
pd <- data.frame(sample_name = sub(basename(cdfs), pattern = ".CDF",
                   replacement = "", fixed = TRUE),
         sample_group = c(rep("KO", 6), rep("WT", 6)),
         stringsAsFactors = FALSE) 
```

```{r}
raw_data <- readMSData(files = cdfs, pdata = new("NAnnotatedDataFrame", pd),
               mode = "onDisk")
```

```{r}
head(rtime(raw_data))
```

```{r}
mzs <- mz(raw_data)

## Split the list by file
mzs_by_file <- split(mzs, f = fromFile(raw_data))

length(mzs_by_file)
```

```{r}
## Get the base peak chromatograms. This reads data from the files.
bpis <- chromatogram(raw_data, aggregationFun = "max")
## Define colors for the two groups
group_colors <- paste0(brewer.pal(3, "Set1")[1:2], "60")
names(group_colors) <- c("KO", "WT")

## Plot all chromatograms.
plot(bpis, col = group_colors[raw_data$sample_group])
```

* bpis[1, i] i代表不同文件（样本） 1-12  返回chromatogram数据
```{r}
bpi_1 <- bpis[1, 1]
head(rtime(bpi_1))
```

* plot中该点的intensity， 本例中是F01.S0001的max intensity
```{r}
head(intensity(bpi_1)) 
```
### EIC

```{r}
## Get the total ion current by file
tc <- split(tic(raw_data), f = fromFile(raw_data))
boxplot(tc, col = group_colors[raw_data$sample_group],
    ylab = "intensity", main = "Total ion current") 
```

## Peak Detection 
### CentWave algorithms ROI
```{r}
## Define the rt and m/z range of the peak area
rtr <- c(2700, 2900)
mzr <- c(334.9, 335.1)
## extract the chromatogram
chr_raw <- chromatogram(raw_data, mz = mzr, rt = rtr)
plot(chr_raw, col = group_colors[chr_raw$sample_group])
```

> Note that Chromatogram objects extracted by the chromatogram method contain an NA value if in a certain scan (i.e. for a specific retention time) **no signal was measured in the respective mz range**. This is reflected by **the lines not being drawn as continuous lines **in the plot above.  

```{r}
raw_data %>%
    filterRt(rt = rtr) %>%
    filterMz(mz = mzr) %>%
    plot(type = "XIC") 
```

```{r}
cwp <- CentWaveParam(peakwidth = c(20, 80), noise = 5000, ppm=5)
xdata <- findChromPeaks(raw_data, param = cwp) 
```

```{r}
head(chromPeaks(xdata))
tail(chromPeaks(xdata))
table(chromPeaks(xdata)[,11])
```

```{r}
summary_fun <- function(z) {
    c(peak_count = nrow(z), rt = quantile(z[, "rtmax"] - z[, "rtmin"]))
}
T <- lapply(split.data.frame(chromPeaks(xdata),
                 f = chromPeaks(xdata)[, "sample"]),
        FUN = summary_fun)
T <- do.call(rbind, T)
rownames(T) <- basename(fileNames(xdata))
pandoc.table(T,
         caption = paste0("Summary statistics on identified chromatographic",
                  " peaks. Shown are number of identified peaks per",
                  " sample and widths/duration of chromatographic ",
                  "peaks.")) 
```

```{r}
plotChromPeaks(xdata, file = 3) 
```

```{r}
plotChromPeakImage(xdata)
```

```{r}
plot(chr_raw, col = group_colors[chr_raw$sample_group], lwd = 2)
highlightChromPeaks(xdata, border = group_colors[chr_raw$sample_group],
            lty = 3, rt = rtr, mz = mzr, type = "rect") 
```

```{r}
plot(chr_raw, col = group_colors[chr_raw$sample_group], lwd = 2)
highlightChromPeaks(xdata, col = group_colors[chr_raw$sample_group],
            lty = 3, rt = rtr, mz = mzr, border = NA,
            type = "polygon") 
```

```{r}
plot(chr_raw, col = group_colors[chr_raw$sample_group], lwd = 2)
highlightChromPeaks(xdata, col = group_colors[chr_raw$sample_group],
            lty = 3, rt = rtr, mz = mzr,
            type = "point") 
```
> Note that we can also specifically extract identified chromatographic peaks for a **selected region by providing the respective m/z and retention time ranges with the mz and rt arguments** in the  `chromPeaks` method.  

```{r}
pander(chromPeaks(xdata, mz = mzr, rt = rtr),
       caption = paste("Identified chromatographic peaks in a selected ",
               "m/z and retention time range.")) 
```

```{r}
## Extract a list of per-sample peak intensities (in log2 scale)
ints <- split(log2(chromPeaks(xdata)[, "into"]),
          f = chromPeaks(xdata)[, "sample"])
boxplot(ints, varwidth = TRUE, col = group_colors[xdata$sample_group],
    ylab = expression(log[2]~intensity), main = "Peak intensities")
grid(nx = NA, ny = NULL) 
```

## Alignment

```{r}
xdata <- adjustRtime(xdata, param = ObiwarpParam(binSize = 0.6)) 
```

```{r}
head(adjustedRtime(xdata))
head(rtime(xdata))
```

```{r}
head(rtime(xdata, adjusted=FALSE))
```

```{r}
## Get the base peak chromatograms.
bpis_adj <- chromatogram(xdata, aggregationFun = "max")
par(mfrow = c(2, 1), mar = c(4.5, 4.2, 1, 0.5))
plot(bpis_adj, col = group_colors[bpis_adj$sample_group])
## Plot also the difference of adjusted to raw retention time.
plotAdjustedRtime(xdata, col = group_colors[xdata$sample_group]) 
```

```{r}
hasAdjustedRtime(xdata)
## Drop the alignment results.
xdata <- dropAdjustedRtime(xdata)

## Does the object have adjusted retention times?
hasAdjustedRtime(xdata) 
```

```{r}
## Correspondence: group peaks across samples.
pdp <- PeakDensityParam(sampleGroups = xdata$sample_group,
            minFraction = 0.8)
xdata <- groupChromPeaks(xdata, param = pdp)

## Now the retention time correction.
pgp <- PeakGroupsParam(minFraction = 0.85)

## Get the peak groups that would be used for alignment.
xdata <- adjustRtime(xdata, param = pgp)
```

```{r}
## Correspondence: group peaks across samples.
pdp <- PeakDensityParam(sampleGroups = xdata$sample_group,
            minFraction = 0.8)
xdata <- groupChromPeaks(xdata, param = pdp)

## Now the retention time correction.
pgp <- PeakGroupsParam(minFraction = 0.85)

## Get the peak groups that would be used for alignment.
xdata <- adjustRtime(xdata, param = pgp)
```

```{r}
## Plot the difference of adjusted to raw retention time.
plotAdjustedRtime(xdata, col = group_colors[xdata$sample_group],
          peakGroupsCol = "grey", peakGroupsPch = 1) 
```
```{r}
par(mfrow = c(2, 1))
## Plot the raw data
plot(chr_raw, col = group_colors[chr_raw$sample_group])

## Extract the chromatogram from the adjusted object
chr_adj <- chromatogram(xdata, rt = rtr, mz = mzr)
plot(chr_adj, col = group_colors[chr_raw$sample_group]) 
```

## Correspondence

```{r}
## Define the mz slice.
mzr <- c(305.05, 305.15)

## Extract and plot the chromatograms
chr_mzr <- chromatogram(xdata, mz = mzr, rt = c(2500, 4000))
par(mfrow = c(3, 1), mar = c(1, 4, 1, 0.5))
cols <- group_colors[chr_mzr$sample_group]
plot(chr_mzr, col = cols, xaxt = "n", xlab = "")
## Highlight the detected peaks in that region.
highlightChromPeaks(xdata, mz = mzr, col = cols, type = "point", pch = 16)
## Define the parameters for the peak density method
pdp <- PeakDensityParam(sampleGroups = xdata$sample_group,
            minFraction = 0.4, bw = 30)
par(mar = c(4, 4, 1, 0.5))
plotChromPeakDensity(xdata, mz = mzr, col = cols, param = pdp,
             pch = 16, xlim = c(2500, 4000))
## Use a different bw
pdp <- PeakDensityParam(sampleGroups = xdata$sample_group,
            minFraction = 0.4, bw = 20)
plotChromPeakDensity(xdata, mz = mzr, col = cols, param = pdp,
             pch = 16, xlim = c(2500, 4000)) 
```

```{r}
## Perform the correspondence
pdp <- PeakDensityParam(sampleGroups = xdata$sample_group,
            minFraction = 0.4, bw = 20)
xdata <- groupChromPeaks(xdata, param = pdp) 
```

The results from the correspondence can be extracted using the `featureDefinitions` method,  

```{r}
featureDefinitions(xdata)
```

```{r}
## Extract the into column for each feature.
head(featureValues(xdata, value = "into"))
```

```{r}
## Filling missing peaks using default settings. Alternatively we could
## pass a FillChromPeaksParam object to the method.
xdata <- fillChromPeaks(xdata)

head(featureValues(xdata, value="mz"))
```
> It should be mentioned that `fillChromPeaks` uses columns "rtmin", "rtmax", "mzmin" and  "mzmax" from the `featureDefinitions` table to define the region from which the signal should be integrated to generate the filled-in peak signal. These values correspond however to the positions of the **peak apex** not the peak boundaries of all chromatographic peaks assigned to a feature. It might be advisable to increase this area in retention time dimension by a constant value appropriate to the average peak width in the experiment. Such a value can be specified with  `fixedRt` of the `FillChromPeaksParam`. If the average peak width in the experiment is 20 seconds, specifying `fixedRt = 10` ensures that the area from which peaks are integrated is at least 20 seconds wide.

```{r}
apply(featureValues(xdata, filled = FALSE), MARGIN = 2,
      FUN = function(z) sum(is.na(z)))
apply(featureValues(xdata), MARGIN = 2,
      FUN = function(z) sum(is.na(z)))
```

```{r}
fcpp <- FillChromPeaksParam(fixedRt = 10)
xdata <- fillChromPeaks(xdata, param = fcpp)
head(featureValues(xdata, value="mz"))
```

```{r}
apply(featureValues(xdata, filled = FALSE), MARGIN = 2,
      FUN = function(z) sum(is.na(z)))
apply(featureValues(xdata), MARGIN = 2,
      FUN = function(z) sum(is.na(z)))
```

> Next we use the `featureSummary` function to get a general per-feature summary that includes the number of samples in which a peak was found or the number of samples in which more than one peak was assigned to the feature. Specifying also sample groups breaks down these summary statistics for each individual sample group.   

```{r}
head(featureSummary(xdata, group = xdata$sample_group))
```

```{r}
feature_chroms <- featureChromatograms(xdata, features = 1:4)

feature_chroms 
par(mfrow = c(1, 2))
plot(feature_chroms[3, ], col = group_colors[feature_chroms$sample_group])
plot(feature_chroms[4, ], col = group_colors[feature_chroms$sample_group]) 
```

```{r}
## Extract the features and log2 transform them
ft_ints <- log2(featureValues(xdata, value = "into"))

## Perform the PCA omitting all features with an NA in any of the
## samples. Also, the intensities are mean centered.
pc <- prcomp(t(na.omit(ft_ints)), center = TRUE)

## Plot the PCA
cols <- group_colors[xdata$sample_group]
pcSummary <- summary(pc)
plot(pc$x[, 1], pc$x[,2], pch = 21, main = "", 
     xlab = paste0("PC1: ", format(pcSummary$importance[2, 1] * 100,
                   digits = 3), " % variance"),
     ylab = paste0("PC2: ", format(pcSummary$importance[2, 2] * 100,
                   digits = 3), " % variance"),
     col = "darkgrey", bg = cols, cex = 2)
grid()
text(pc$x[, 1], pc$x[,2], labels = xdata$sample_name, col = "darkgrey",
     pos = 3, cex = 2)
```

