---
title: "MiddlePlot"
output: html_notebook
---

# scatterPlot
添加拟合曲线
```{r}
attach(mtcars)
plot(wt, mpg, 
     main='Basic Scatter',
     xlab='Car Weight',
     ylab='Miles Per Gallon')
abline(lm(mpg~wt), col='red', lwd=2, lty=1)
lines(lowess(wt, mpg), col='blue', lwd=2, lty=2)
```
## scatterplot() in car
```{r}
library(car)
scatterplot(mpg ~ wt | cyl, data=mtcars, lwd=2, span=.75,
            legend.plot=TRUE, id.method='identify', 
            #labels=row.names(mtcars),
            boxplots='xy')
```

## scatter matrix
```{r}
pairs(~mpg+disp+drat+wt, data=mtcars, 
      main='Basic Scatter Plot Matrix')
```
```{r}
library(car)
scatterplotMatrix(~mpg+disp+drat+wt, data=mtcars,
                  main='Scatter Plot Matrix via car package')
```

* even more: cpars() in glus package; pairs2() in TeachingDemos; xysplom() in HH; kepairs() in ResourceSeletion;pairs.mod() in SMPracticals

```{r}

```

## high density scatterplot

```{r}
set.seed(1234)
n <- 10000
c1 <- matrix(rnorm(n, mean=0, sd=.5), ncol=2)
c2 <- matrix(rnorm(n, mean=3, sd=2), ncol=2)
mydata <- rbind(c1, c2)
mydata <- as.data.frame(mydata)
names(mydata) <- c('x', 'y')
with(mydata, 
     plot(x, y, pch=19, main='Scatter Plot'))
```
hard to decide the density  

```{r}
with(mydata, smoothScatter(x, y))
```

## hexbin()
```{r}
library(hexbin)
with(mydata, {
  bin <- hexbin(x, y, xbins=50)
  plot(bin, main='Hexagonal Binning with 10,000 observations')
})
```
get more examples by typing `?smoothScatter` or `?ipairs`  

## 3d

```{r}
library(scatterplot3d)
attach(mtcars)
scatterplot3d(wt, disp, mpg,
              pch=16, highlight.3d=TRUE,
              type='h')
```

add regression plane  
```{r}
library(scatterplot3d)
attach(mtcars)
s3d <- scatterplot3d(wt, disp, mpg,
              pch=16, highlight.3d=TRUE,
              type='h')
fit <- lm(mpg ~ wt+disp)
s3d$plane3d(fit)
```
## rotatable scattplot3d
```{r}
library(rgl)
attach(mtcars)
plot3d(wt, disp, mpg, col='red', size=5)
detach(mtcars)
```

```{r}
library(car)
with(mtcars, scatter3d(wt, disp, mpg))
```

## bubble plot
```{r}
attach(mtcars)
r <- sqrt(disp/pi)
symbols(wt, mpg, circle=r, inches=.30,
        fg='white', bg='lightblue')
text(wt, mpg, rownames(mtcars), cex=.6)
```

# line chart
```{r}
opar <- par(no.readonly = TRUE)
par(mfrow=c(1,2))
t1 <- subset(Orange, Tree==1)
plot(t1$age, t1$circumference)
plot(t1$age, t1$circumference, type='b')
```

```{r}
Orange$Tree <- as.numeric(Orange$Tree)
ntrees <- max(Orange$Tree)

xrange <- range(Orange$age)
yrange <- range(Orange$circumference)

plot(xrange, yrange, type='n', xlab='Age (days)', ylab='Circumference (mm)')

colors <- rainbow(ntrees)
linetype <- c(1:ntrees)
plotchar <- seq(18, 18+ntrees, 1)
for (i in 1:ntrees) {
  tree <- subset(Orange, Tree==i)
  lines(tree$age, tree$circumference,
        type='b', lwd=2, lty=linetype[i],
        col=colors[i], pch=plotchar[i])
}

title("Tree Growth", 'example of line chart')
legend(xrange[1], yrange[2],
       1:ntrees, cex=.8, col=colors, pch=plotchar, lty=linetype, title='Tree')
```

# correlation plot
```{r}
options(digits=2)
cor(mtcars)
```

```{r}
library(corrgram)
corrgram(mtcars, order=TRUE, lower.panel=panel.shade,
         upper.panel=panel.pie, text.panel=panel.txt,
         main='Corrgram of mtcars intercorrelations')
```

```{r}
library(corrgram)
corrgram(mtcars, order=TRUE, lower.panel=panel.ellipse,
         upper.panel=panel.pts, text.panel=panel.txt,
         diag.panel=panel.minmax,
         main='Corrgram of mtcars data using scatter plots and ellipse')
```

```{r}
library(corrgram)
corrgram(mtcars, lower.panel = panel.shade,
         upper.panel = NULL, text.panel=panel.txt,
         main='Car Mileage Data')
```

```{r}
library(corrgram)
cols <- colorRampPalette(c('darkgoldenrod4', 'burlywood1', 'darkkhaki', 'darkgreen'))
corrgram(mtcars, col.regions=cols,
         lower.panel = panel.shade,
         upper.panel = panel.conf, text.panel=panel.txt,
         main='Car Mileage Data')
```

# Mosaic plot

```{r}
library(vcd)
mosaic(Titanic, shade=TRUE, legend=TRUE)
```

```{r}
library(vcd)
mosaic(~Class+Sex+Age+Survived, data=Titanic, shade=TRUE, legend=TRUE)
```

