---
title: "Regression"
output: html_notebook
---

### OLS
#### Simple Linear Regression
```{r}
fit <- lm(weight ~ height, data=women)
summary(fit)
```
```{r}
residuals(fit)
```

```{r}
plot(fit)
```
#### 多项式回归
```{r}
fit2 <- lm(weight ~ height + I(height ^ 2), data=women)
summary(fit2)
plot(women$height, women$weight, 
     xlab='Height', ylab='Weight')
lines(women$height, fitted(fit2))
```
```{r}
library(car)

car::scatterplot(weight ~ height, data=women, 
            pch=19,  
            main='Women Age 30-39', 
            )
```

#### 多元线性回归
```{r}
states <- as.data.frame(state.x77[, c('Murder', 'Population', 'Illiteracy', 'Income', 'Frost')])
```

You'd better check the correlation between varabiles before implement of model fitting.
```{r}
cor(states)
```
```{r}
scatterplotMatrix(states, 
                  main='Scatter Matrix')
```
```{r}
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)
summary(fit)
```

#### 有交互项
```{r}
fit <- lm(mpg ~ hp + wt + hp:wt, data=mtcars)
summary(fit)
plot(effect('hp:wt', fit, list=(wt=c(2.2, 3.2, 4.2))), multiline=TRUE)
```

### 回归诊断
```{r}
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)
confint(fit)
```

```{r}
plot(fit)
```

```{r}
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)
par(mfrow=c(2,2))
plot(fit)
```
#### 正态性
```{r}
library(car)
qqPlot(fit, labels=row.names(states), id.method='identify', simulate=TRUE, main='Q-Q Plot')
```

```{r}
residplot <- function(fit, nbreaks=10) {
  z <- rstudent(fit)
  hist(z, breaks=nbreaks, freq=FALSE, 
       xlab='Studentized Residual', 
       main='Distribution of Errors')
  rug(jitter(z), col='brown')
  curve(dnorm(x, mean=mean(z), sd=sd(z)), 
        add=TRUE, col='blue', lwd=2)
  lines(density(z)$x, density(z)$y, 
        col='red', lwd=2, lty=2)
  legend('topright', 
         legend=c('Normal Curve', 'Kernel Density Curve'), 
         lty=1:2, col=c('blue', 'red'), cex=.7)
}

residplot(fit)
```

#### 误差独立性
Durbin-Watson test
```{r}
durbinWatsonTest(fit)
```
* p is not significant，indicating there is no self-autocorrelation

#### 线性
成分残差图 (component plus residual plot, or partial residual plot)

```{r}
crPlots(fit)
```

#### 同方差性

```{r}
ncvTest(fit)
spreadLevelPlot(fit)
```

### Comprehensive Validation
`gvlma()` in gvlma

```{r}
library(gvlma)
gvmodel <- gvlma(fit)
summary(gvmodel)
```

### Multicollinearity 多重共线性
* 统计量VIF(Variance Inflation Factor) 方差膨胀因子

```{r}
sqrt(vif(fit))
sqrt(vif(fit)) > 2
```

### 异常值观测

#### 离群点